---
title: "Project Notebook"
output: html_notebook
---
This file is part of the Project_Starter-IKEA_daring_coua_of_expression distribution 
(https://git.wur.nl/geoscripting-2023-january/staff/project/Project_Starter-IKEA_daring_coua_of_expression).
 Copyright (c) 2023 IKEA_daring_coua_of_expression.

This program is free software: you can redistribute it and/or modify  it under the terms of the GNU General Public License as published by  the Free Software Foundation, version 3.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the GNU General Public License for more details.
You should have received a copy of the GNU General Public License along with this program. If not, see <http://www.gnu.org/licenses/>.

# Geoscripting 2023 Final Project

## Challenge 2 - Tree Species Classification using UAV Imagery

## Team: IKEA Coua

### Step 1 - Download data and create directories

#### ATTENTION: After creating the directories, we need to manually download data from Google Drive and save them in the 'data' folder before proceeding further.

Run this line first
```{r}
install.packages("rmarkdown", dependencies=TRUE)
```


```{r}
# Import function scripts
source("R/make_directories.R")

# Use the function to download data and create directories
make_directories()
```

#### Preview the data

```{r}

# Install and load terra
if(!"terra" %in% installed.packages()) {install.packages("terra")}
library(terra)

# Create SpatRaster objects with terra
chm <- terra::rast("data/CHM.tif")
multispectral <- terra::rast("data/Multispectral.tif")
thermal <- terra::rast("data/Thermal.tif")

# Plot the SpatRasters
terra::plot(chm)
terra::plotRGB(multispectral, 1, 2, 3, stretch = "lin")
terra::plotRGB(thermal)
```

### Step 2 - Find individual tree crowns

```{r}
# Install and load packages
if(!"sf" %in% installed.packages()) {install.packages("sf")}
if (!"itcSegment" %in% installed.packages()) {install.packages("itcSegment")}
library(sf)
library(itcSegment)

# if (!"terra" %in% installed.packages()) {install.packages("terra")}
# library(terra)

# Import function scripts
source('R/SegmentTreeCrowns.R')

# Find the path to CHM.tif
## Update it to search masked at the end
chm_path <- list.files(path = "data/", pattern = glob2rx('CHM.tif'), full.names = TRUE)

# Use the function to create the masked CHM.tif
tree_polygons <- SegmentTreeCrowns(chm_path)
```

#### Plot tree crowns on top of thermal image

```{r}
terra::plotRGB(thermal)
terra::plot(st_geometry(tree_polygons), add = TRUE, col = 'red')
```

### Step 3 - Combine mean of multispectral and thermal value with polygons to make training data

```{r}
# Install and load packages
if (!"dplyr" %in% installed.packages()) {install.packages("dplyr")}
library(dplyr)

# if (!"terra" %in% installed.packages()) {install.packages("terra")}
# if (!"sf" %in% installed.packages()) {install.packages("sf")}
# library(terra)
# library(sf)

# Import function scripts
source('R/make_training_data.R')

# Generate dataframe that includes all band information and tree height

tree_polygons_path <- "data/Tree_Segmentation_Polygons.shp"

training_data <- make_training_data(tree_polygons_path)
```

#### Investigate the structure of training_data

```{r}
head(training_data)
```

### Step 4.1 - Cluster the polygons via KMeans with python 

```{r}
# This step will be done in Python. The result plot is saved in folder 'output'.

# Install and load packages to use python functions in R
if (!require("reticulate", character.only = TRUE)) {
  install.packages("reticulate")
  require("reticulate", character.only = TRUE)
}
library(reticulate)

# Use the conda environment "clustering" that was created when running ikea_coua.sh
use_condaenv("clustering")

# Run the python scripts
py_run_file("python/functions/cluster_polygons_KMeans.py")

# Based on the derived graph, the best K equals to 5
K <- 5

####---------------------------------------------------------------
```

### Step 4.2 - Cluster the polygons via KMeans with R

```{r}
# Install and load packages
if (!"cluster" %in% installed.packages()) {install.packages("cluster")}
library(cluster)

# if(!"terra" %in% installed.packages()){install.packages("terra")}
# if(!"sf" %in% installed.packages()){install.packages("sf")}
# library(terra)
# library(sf)

# Import function scripts
source("R/ClusterPolygons.R")

# Use the function to cluster the polygons with the derived best K

training_data_path <- "data/training_data.shp"
ClusterPolygons(training_data_path, K = K, withThermal = FALSE)
ClusterPolygons(training_data_path, K = K, withThermal = TRUE)
```

### Step 4.3 Clustering with GMM and calculate silhouette values

```{r}
py_run_file("python/GMM_clustering_and_silhouette_values_calculation.py")
```



#### Step 5 - Visualize the clusters

```{r}
# Install and load packages
if (!"ggplot2" %in% installed.packages()) {install.packages("ggplot2")}
library(ggplot2)

# Import function scripts
source("R/VisualiseClusters.R")

# Visualize the clusters
VisualiseClusters(path="output/Python_Polygons_clustered_Kmeans_with_Thermal.shp")
VisualiseClusters(path="output/Python_Polygons_clustered_Kmeans_without_Thermal.shp")
VisualiseClusters(path="output/Python_Polygons_clustered_GMM_with_Thermal.shp")
VisualiseClusters(path="output/Python_Polygons_clustered_GMM_without_Thermal.shp")
VisualiseClusters(path="output/Python_Polygons_clustered_Kmeans_with_Thermal.shp")
VisualiseClusters(path="output/Python_Polygons_clustered_Kmeans_without_Thermal.shp")


# Install and load packages
if(!"magick" %in% installed.packages()){install.packages("magick")}
library(magick)

# View the png files
species_with_thermal <- magick::image_read('output/Tree_Species_Clusters_KMeans_without_Thermal.png')
species_without_thermal <- magick::image_read('output/Tree_Species_Clusters_KMeans_without_Thermal.png')
plot(species_with_thermal)
plot(species_without_thermal)
```
